---
title: "Forecasting Energy Demand in R"
format: 
  revealjs:
    math: true
    theme: solarized
editor: visual
execute: 
  allow: true
bibliography: references.bib
rsconnect::writeManifest()
---

```{r}
#| include: false
#| results: hide
source("Energy Demand Spain.R", print.eval = FALSE)
```

# My Goals

-   Explore different time series forecasting techniques in R

-   See how accurate a forecast I can develop

-   Understand the challenges with forecasting this specific dataset

-   I published a series of LinkedIn posts discussing this project

    -   [My profile](https://www.linkedin.com/in/zachary-nabavian/)

    -   [My summary post](https://www.linkedin.com/pulse/lessons-learned-from-my-energy-demand-forecasting-project-nabavian-wnmhe/?trackingId=1JOoGdvNSM%2BKmxpCE6WUwA%3D%3D)

    -   [My GitHub](https://github.com/Zach-Nabavian/Energy_Demand)

# About the Dataset

-   Contains four years of total hourly electricity demand from Spain, from 2015 to 2018

-   Also contains weather data over the same period

-   [The dataset's Kaggle page](https://www.kaggle.com/datasets/nicholasjhana/energy-consumption-generation-prices-and-weather)

# Understanding Seasonality of the Data

-   Energy demand is unique in that it has *multiple kinds of seasonality*

Demand varies by the following:

-   By hour of the day

-   by the day of the week

-   and by the month of the year

# Hourly Seasonality

```{r}
hour_box_plot
```

# Weekly Seasonality

```{r}
week_box_plot
```

# Monthly Seasonality

```{r}
month_box_plot
```

# The Challenge

-   My initial forecasting models could only handle one kind of seasonality at a time

-   My options included:

    -   Just build a model with hourly seasonality, since that's the strongest

    -   Try to incorporate all three by adding fourier terms

    -   I did eventually find a better solution

# Short Term Forecast

-   I initially worked on short term forecasts to avoid the problem of multiple seasonalities

-   The goal of these forecasts was to forecast demand for February 2015 using the data from January 2015

-   The simple ARIMA model did best

# Short Term Forecast Accuracy

```{r}
short_term_table
```

# Issues With More Complex Models

-   These models are harder to explain

-   They take much longer to run

-   I tried to include fourier terms to account for multiple seasonalities, but this model takes much longer to run

-   At one point, I tried to use a machine learning model, but I ultimately wanted to use models that didn't take so long to run

# Extending to Longer Term

-   My next goal was to forecast the whole year 2018 using the data from 2015 to 2017

-   My initial forecast models used for the short term forecast were all less accurate

-   These models were all surprisingly biased in the same direction

-   I did eventually find a more accurate forecast

# Long Term Forecast Accuracy

```{r}
long_term_measures_table
```

# Finding Academic Literature on This Subject

-   I wanted to see how other approached forecasting energy demand

-   I found [this academic journal article](https://www.sciencedirect.com/science/article/abs/pii/S0169207015001557)[^1]

-   The forecast model suggested uses a cubic regression model incorporating temperature values

-   It does not incorporated prior values of energy demand, unlike my other forecasts

-   The regression also includes terms for hour, weekday, and month, solving the issue of multiple seasonalities

[^1]: [@wang2016]

# The Regression Model

$$
\begin{align}
f(T_t) &= B_6Tt + B_7T_t^2 + B_8T_t^3 \\ 
&\quad + B_9T_tM_t + B_{10}T_t^2M_t + B_{11}T_t^3M_t \\ 
&\quad + B_{12}T_tH_t + B_{13}T_t^2H_t + B_{14}T_t^3H_t
\end{align}
$$

# Why Use a Cubic Regression?

```{r}
temp_energy_scatter
```

# Comparing Performance

```{r}
long_term_measures_table
```

# Issues With The Regression Model

-   If we want to put this model into practice, we won't be able to use real-time temperature data if we want to forecast future demand

-   Therefore, I believe this model in practice would need to be combined with a weather forecast model

-   Including recency terms don't seem to improve the model

-   This model is not meant to be some revolutionary new forecast that's better than what the industry is already doing

# Lessons Learned

-   Understanding what experts in the field are doing is more important than using the most cutting edge technology

-   I used the book [*Forecasting: Principles and Practice*](https://otexts.com/fpp3/bibliography.html)to help learn how to build these models in R

-   Posting on LinkedIn helped me understand what people are interested in seeing in these kinds of presentations
